\documentclass[12pt,a4paper]{article}

%% ============================================================
%% Packages
%% ============================================================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{cleveref}

%% ============================================================
%% Theorem environments
%% ============================================================
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

%% ============================================================
%% Custom commands
%% ============================================================
\newcommand{\Psys}{P_{\mathrm{sys}}}
\newcommand{\Peff}{P_{\mathrm{eff}}}
\newcommand{\Fsys}{F_{\mathrm{sys}}}
\newcommand{\Rel}{\mathrm{Rel}}
\DeclareMathOperator*{\argmax}{arg\,max}

%% ============================================================
%% Title
%% ============================================================
\title{%
  Probabilistic Scaling Laws for Agent Systems:\\[4pt]
  \large From Multiplication Laws to Unified Reliability Polynomials%
}
\author{%
  Agent O \and Agent C \and Orchestrator%
}
\date{March 2026}

\begin{document}
\maketitle

%% ============================================================
%% Abstract
%% ============================================================
\begin{abstract}
We develop a rigorous probabilistic framework for the scaling behavior of
multi-agent systems by modeling each agent as a Bernoulli random variable and
deriving composition laws that govern system-level performance.  Our results
span five regimes: (1)~independent sequential pipelines, where system
reliability decays exponentially with a universal throughput constant
$\Psys(n^*) = e^{-1}$; (2)~dependent pipelines, analyzed as absorbing Markov
chains with coordination gain determined by the geometric mean of coordination
multipliers; (3)~parallel ensembles, exhibiting diminishing returns governed by
the Condorcet jury theorem; (4)~hybrid series-parallel topologies, optimized
via a marginal equalization principle with a unique efficiency maximum; and
(5)~capability saturation, characterized by architecture-specific thresholds
and a phase diagram in capability--cost space.  All five regimes are unified as
special cases of a reliability polynomial on the task as a Directed Acyclic Graph (DAG).  Our theoretical
predictions are quantitatively consistent with the empirical findings of Kim
et~al.\ (2025), including the tool-coordination tradeoff ($R^2 = 0.267$),
capability saturation at $\sim\!45\%$ baseline, centralized coordination
yielding $+81\%$ on structured tasks, independent coordination yielding
$-70\%$ on sequential planning, and task-contingent optimal architecture with
$87\%$ prediction accuracy.
\end{abstract}

\tableofcontents
\newpage

%% ============================================================
%% Section 0: Introduction
%% ============================================================
\section{Introduction}
\label{sec:introduction}

The question of how multi-agent systems scale --- whether adding agents improves
or degrades overall performance --- is fundamental to the design of AI systems
that coordinate multiple language model agents.  Empirical work by Kim
et~al.~(2025) has identified several striking regularities: error amplification
in sequential pipelines, diminishing returns in parallel ensembles, a
capability saturation threshold near $45\%$, and task-contingent optimal
architectures predictable from task topology features.  These findings call
for a formal mathematical framework that can explain and predict scaling
behavior from first principles.

This paper provides such a framework.  We model each agent as a Bernoulli
random variable and derive the probability laws governing system success
under five increasingly general task topologies:

\begin{enumerate}[label=\textbf{Regime \arabic*:}, leftmargin=*]
  \item \textbf{Independent Sequential Pipelines} (\Cref{sec:regime1}).
    The multiplication law yields exponential decay $\Psys = p^n$, a universal
    throughput constant $e^{-1}$, and the result that adding agents always
    degrades effective performance for any coordination cost $c > 0$.

  \item \textbf{Dependent Sequential Pipelines} (\Cref{sec:regime2}).
    The chain rule decomposition, Markov pipeline analysis, and centralized
    validation theory establish when coordination improves reliability and
    when it amplifies errors.

  \item \textbf{Parallel Ensembles} (\Cref{sec:regime3}).
    Diminishing returns follow a geometric sequence with ratio $(1-p)$, the
    Condorcet jury theorem governs majority voting, and parallel architectures
    strictly dominate sequential ones for $n \geq 2$.

  \item \textbf{Hybrid Series-Parallel Topologies} (\Cref{sec:regime4}).
    Recursive composition on series-parallel DAGs, the marginal equalization
    principle for optimal resource allocation, and a unique efficiency maximum
    per topology class.

  \item \textbf{Capability Saturation and the Unified Scaling Law}
    (\Cref{sec:regime5}).
    Architecture-specific saturation thresholds, a phase diagram in $(p,c)$-space,
    and the reliability polynomial as the unifying mathematical object
    encompassing all regimes.
\end{enumerate}

\subsection{Notation}

Throughout, $A_i$ denotes an agent modeled as a Bernoulli random variable with
success probability $p_i = P(A_i = 1)$.  We write $p$ for the uniform
capability when all agents share the same success probability, $q = 1-p$ for
the failure probability, and $c$ for the per-link (or per-agent) coordination
cost.  The system success probability is $\Psys$ and the effective system
probability (including cost) is $\Peff$.  The error amplification factor is
$\alpha = p(1-c)$.

\subsection{Relationship to Prior Work}

Our framework is motivated by the empirical findings of Kim et~al.\ (2025),
\emph{Towards a Science of Scaling Agent Systems}, which studied multi-agent
coordination across diverse benchmarks.  We provide theoretical foundations for
their observations, deriving each empirical pattern from the probabilistic
structure of the task topology.  The correspondence between our predictions and
their findings is summarized in \Cref{sec:regime5} (Theorem~\ref{thm:empirical-correspondence}).

\newpage

%% ============================================================
%% Section 1: Definitions
%% ============================================================
\input{definitions}

\newpage

%% ============================================================
%% Section 2: Regime 1
%% ============================================================
\input{regime_1_independent}

\newpage

%% ============================================================
%% Section 3: Regime 2
%% ============================================================
\input{regime_2_dependent}

\newpage

%% ============================================================
%% Section 4: Regime 3
%% ============================================================
\input{regime_3_parallel}

\newpage

%% ============================================================
%% Section 5: Regime 4
%% ============================================================
\input{regime_4_hybrid}

\newpage

%% ============================================================
%% Section 6: Regime 5
%% ============================================================
\input{regime_5_saturation}

\newpage

%% ============================================================
%% Section 7: Discussion
%% ============================================================
\section{Discussion}
\label{sec:discussion}

\subsection{Summary of Principal Results}

We have established a hierarchy of scaling laws for multi-agent systems, each
building on the previous:

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{>{\bfseries}l l l}
\toprule
Regime & Key Formula & Central Result \\
\midrule
1. Independent &
  $\Psys = p^n$ &
  Exponential decay; $\Psys(n^*) = e^{-1}$ universally \\
2. Dependent &
  $\Psys = p_1 \cdot r^{n-1}$ &
  Coordination helps iff $r > p$ \\
3. Parallel &
  $\Psys = 1-(1-p)^n$ &
  Diminishing returns; Condorcet at $p=1/2$ \\
4. Hybrid &
  Recursive SP &
  Minimize series depth; equalize parallel width \\
5. Unified &
  $\Rel(G; \mathbf{p}, \mathbf{c})$ &
  All regimes are special cases \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Design Implications}

The theoretical framework yields concrete design principles:

\begin{enumerate}
  \item \textbf{Sequential pipelines are generically harmful.}
    For any $c > 0$, the critical threshold $p^* = 1/(1-c) > 1$ is
    unattainable, so adding sequential agents always degrades performance
    (Theorem~\ref{thm:critical-threshold}).  Sequential stages should be
    minimized and supplemented with parallel redundancy
    (Theorem~\ref{thm:r4-optimal-allocation}).

  \item \textbf{Architecture selection dominates capability improvement.}
    At $p = 0.45$, switching from sequential to parallel provides a
    $50\times$ improvement at $n = 5$, while improving $p$ from $0.45$ to
    $0.50$ gives only $11\%$ for a single agent.  The topology of coordination
    matters far more than the capability of individual agents.

  \item \textbf{The Condorcet threshold is a hard boundary.}
    Majority voting requires $p > 1/2$ to be beneficial
    (Theorem~\ref{thm:majority-threshold}).  For $p < 1/2$, parallel-any
    (at-least-one) is the only viable multi-agent strategy.

  \item \textbf{Every architecture has a finite optimal team size.}
    For any coordination cost $c > 0$, there exists a finite $n^*$ beyond
    which adding agents is detrimental
    (Theorems~\ref{thm:cost-dominance-seq}--\ref{thm:cost-dominance-maj}).

  \item \textbf{The $\sim\!45\%$ saturation threshold is architecture-dependent.}
    It reflects the capability at which most standard coordination strategies
    (sequential, majority) fail, leaving only parallel-any viable.  The
    phase diagram (Proposition~\ref{prop:phase-diagram}) makes this precise.
\end{enumerate}

\subsection{Assumptions and Limitations}

Our framework rests on several simplifying assumptions:

\begin{enumerate}
  \item \textbf{Binary outcomes.}  Agents produce binary success/failure.
    A richer model with graded quality would require continuous random
    variables and information-theoretic cost measures.

  \item \textbf{Uniform capability.}  Most results assume $p_i = p$.
    The AM--GM bound (Lemma~\ref{lem:am-gm}) shows this is the
    best-case scenario; heterogeneous agents are strictly worse.

  \item \textbf{Independence within parallel stages.}  We assume
    parallel agents fail independently.  Correlated failures (e.g., from
    shared training data) would reduce the effectiveness of parallel
    redundancy.

  \item \textbf{Series-parallel structure.}  The recursive composition
    theorem requires SP graphs.  General DAGs require the reliability
    polynomial, which is $\#P$-hard to compute exactly
    (Theorem~\ref{thm:r4-complexity}).

  \item \textbf{Static architecture.}  We do not model adaptive
    architectures that dynamically reconfigure based on intermediate
    results.
\end{enumerate}

\subsection{Open Problems}

\begin{enumerate}
  \item \textbf{Continuous-quality agents.}  Extend the binary model to
    agents whose outputs have a quality distribution.  The natural
    framework is information-theoretic: coordination cost becomes mutual
    information between agent outputs.

  \item \textbf{Correlated parallel agents.}  Quantify how shared
    failure modes reduce parallel redundancy gains.  The FKG inequality
    provides a starting point (Theorem~\ref{thm:r4-sp-bounds}).

  \item \textbf{Adaptive architectures.}  Model systems that dynamically
    switch topology based on observed intermediate results.  This
    requires a sequential decision-theoretic framework.

  \item \textbf{Tighter general-DAG bounds.}  Close the gap between the
    SP bounds and exact reliability for non-SP DAGs (cf.\ the Wheatstone
    bridge example in Remark~\ref{rmk:r4-sp-tightness}).

  \item \textbf{Heterogeneous cost structures.}  Allow different
    coordination costs for different links and agents, moving beyond
    the uniform-cost model.
\end{enumerate}

%% ============================================================
%% Section 8: Conclusion
%% ============================================================
\section{Conclusion}
\label{sec:conclusion}

We have developed a complete probabilistic theory of scaling in multi-agent
systems.  The central mathematical object is the \emph{reliability polynomial}
$\Rel(G; \mathbf{p}, \mathbf{c})$ on the task DAG, which unifies five
distinct scaling regimes as special cases.  Our framework provides:

\begin{itemize}
  \item \textbf{Explanatory power}: all six empirical findings of Kim
    et~al.\ (2025) have quantitative theoretical explanations.
  \item \textbf{Predictive power}: the architecture selection function
    (Theorem~\ref{thm:arch-selection}) is a deterministic function of
    three observable task parameters.
  \item \textbf{Design guidance}: concrete principles for when to use
    sequential vs.\ parallel vs.\ hybrid architectures, and how to
    allocate resources across stages.
\end{itemize}

The key insight is that \emph{topology dominates capability}: the
structure of agent coordination matters far more than individual agent
performance.  A well-chosen architecture can yield orders-of-magnitude
improvement over a poorly chosen one, even with identical agents.

\end{document}
